---
title: "Intro to Regression Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome

In this workshop, We will explore the basic regression analysis in R. The datasets consist of the `spotify_lyrics` data modified from [the kaggle dataset](https://www.kaggle.com/zaheenhamidani/ultimate-spotify-tracks-db) and the `spotify` data comprised of the musical attributes of weekly [Billboard Hot 100](https://www.billboard.com/charts/billboard-200/) songs from February 2019 to February 2021. Both of the datasets were mined using the [spotifyr](https://www.rcharlie.com/spotifyr/) interface to the [Spotify Web API](https://developer.spotify.com/documentation/web-api/)*. We will have to load the following packages in this workshop: `tidyverse`,`GGally`,`leaps`. 

Learning objectives: 

- Exploratory data analysis

- Fit regression models with the `lm` function 

- Automatic model selection (package `leaps`)

- Prediction

- Interaction terms 

Along the way, you will master the core functions for running linear regression in R and be able to interpret the regression result generated in R: 

- `pairs()` and `ggpairs()`, which let you explore the relationships among variables 

- `glimpse` and `summary()`, which let you understand the structure of the dataset and descriptive statistics of the variables. 

- `%>%`, which organizes your code into reader-friendly "pipes" 

- `step`, which helps with automatic model selection 

First, let's load the `tidyverse` suite of packages and the `here` package. The `here` package simplifies the way of reproducing file paths. 

```{r load-package, warning = FALSE,message= FALSE}
library(tidyverse)
library(here)
library(GGally)
library(leaps)
```

## Explore Spotify data
We'll first analyze a Spotify dataset and the variables are:

- energy: Energy is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy.

- loudness: 

- tempo:

- danceability: 

- lyrics:

More details about the variable can be found at 

```{r spotify, message = FALSE}
spo <- read_csv("https://raw.githubusercontent.com/YuxiaoLuo/r_analysis_dri_2022/main/data/spotify_lyrics.csv")
glimpse(spo)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.











